### 2018-05-11 doc/paper.tex
keywords

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{subfig}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{hyperref}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}

\usepackage{url} % not crucial - just used below for the URL

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\newenvironment{definition}[1]% environment name 
{% begin code 
  \par\vspace{.75\baselineskip}\noindent 
  \textbf{Definition (#1)}\begin{itshape}% 
  \par\vspace{.5\baselineskip}\noindent\ignorespaces 
}% 
{% end code 
  \end{itshape}\ignorespacesafterend 
}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf Adaption of the Chumbley Score to matching of bullet striation marks}

  \author{
        Ganesh Krishnan \thanks{The authors gratefully acknowledge \ldots{}} \\
    Department of Statistics, Iowa State University\\
     and \\     Heike Hofmann \\
    Department of Statistics and CSAFE, Iowa State University\\
      }
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Adaption of the Chumbley Score to matching of bullet striation marks}
  \end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}

\end{abstract}

\noindent%
{\it Keywords:} 3 to 6 keywords, that do not appear in the title
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

\newcommand{\hh}[1]{{\textcolor{orange}{#1}}}
\newcommand{\gk}[1]{{\textcolor{blue}{#1}}}
\newcommand{\cited}[1]{{\textcolor{red}{#1}}}
\setlength\parindent{0pt}

\tableofcontents
\newpage

\hypertarget{introduction-and-background}{%
\section{Introduction and
Background}\label{introduction-and-background}}

\hypertarget{motivation}{%
\subsection{Motivation}\label{motivation}}

Same source analyses are a major part of an Forensic Toolmark Examiner's
job. In current practice examiners make these comparisons by visual
inspection under a comparison microscope and come to one of the
following four conclusions: identification, inconclusive, elimination or
unsuitable for examination\textasciitilde{}\citep{afte-toolmarks1998}.
These conclusions are made on the basis of ``unique surface contours''
of the two toolmarks being in ``sufficient agreement''
\citep{afte-toolmarks1998}. AFTE describes the term ``sufficient
agreement'' as the possibility of another tool producing the markings
under comparison, as practically impossible \citep{afte-toolmarks1998}.
Potential subject bias in the assessment as well as the lack of
specified error rates are the main points of criticisms first raised by
the National Research Council in 2009 \citep{NAS:2009} and later
emphasized further by the President's Council of Advisors on Science and
Technology \citep{pcast2016}.

Technological advances, such as profilometers and confocal microscopy
allow to measure 3D surfaces in a high-resolution digitized form. This
technology has become more accessible over the last decade, and has made
its way into topological images of ballistics evidence, such as bullet
lands and breech faces
\citep{DeKinder1, DeKinder2, Bachrach1, vorburger2016}. Digitized images
of 3D surfaces of form the data basis of statistical analysis of
toolmarks. A statistical approach based on data removes both
subjectivity from the assessment and allows a quantification of error
rates for both false positive and false negative identifications.

{\textcolor{orange}{In the next page and a half it is easy to lose the red line. It might help to include a table with an overview.
The table should include the reference to the paper, the data used, the statistical method and the associated error rates.}}
{\textcolor{blue}{table included}}

Various toolmarks have been studied in the literature:
\citet{manytoolmarks1} and \citet{chumbley} have been analyzing
screwdriver marks digitized using a profilometer; \citet{manytoolmarks2}
have investigated 3D marks from screwdriver, tongue and groove pliers
captured using a confocal microscope; \citet{afte-chumbley} have been
investigated digitized marks from slip-joint pliers generated by a
surface profilometer.

{\textcolor{orange}{We need an additional sentence here to get from the data to the statistical methods ... }}
{\textcolor{blue}{Analysis of these digitized markings require the use of statistical methods which can quantify the scientific mechanism of comparing markings and serve as metrics for correctly identifying the error rate.}}

\citet{manytoolmarks2} define a relative distance metric and use it as
similarity measure between two toolmarks. \citet{manytoolmarks1} extract
many small segments in the markings of two toolmarks and compare
similiarity using a maximum pearson correlation coefficient. The
Chumbley scoring method, first introduced by \citet{chumbley}, uses a
similar but more extensive framework based on a Mann-Whitney U test of
the resulting correlation coefficients. This approach is
non-deterministic, because segments are chosen randomly. \citep{hadler}
make the score deterministic for each pair of toolmarks by choosing
segments for comparison systematically. This approach also ensures
independence between segements of striae. In this paper, we are
investigating the applicability of the Chumbley scoring method by
\citet{hadler} to assess striation marks on bullet lands for same-source
identification.

\begin{table}[ht]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{rlrrr}
  \hline
Research paper & Method & Data Source & False Positives & False Negatives\\
  \hline
 \textbf{\citet{manytoolmarks1}} & Maximized & Screwdrivers & & \\ & Correlation &  & - & - \\ \hline
 \textbf{\citet{chumbley}} & Randomized & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 2.3\% & 8.9\%  \\ \hline
  \textbf{\citet{afte-chumbley}} & Randomized & Slip-joint  &  & \\ & Chumbley Score &  & - & - \\ \hline 
  \textbf{\citet{hadler}} & Deterministic & Screwdrivers & & \\ (Same-Surface Same-Angle) & Chumbley Score & & 0\% & 6\%  \\ \hline
  \textbf{\citet{manytoolmarks2}} & Similarity Measure & Screwdrivers & & \\ (Different Surfaces-same angle) & Relative Distance Metric & & 5.9\% &  9.4\% \\ (Same Surfaces-same angle)&  & & 0.22\% &  0\% \\
   \hline
\end{tabular}}
\caption{\label{tab:toolmarks-ER} Error Rates for Toolmarks using  variations of the chumbley score method}
\end{table}

Striation marks on bullets are made by impurities in the barrel. As the
bullet travels through the barrel, these imperfections leave
``scratches'' on the bullet surface. Typically, only striation marks in
the land engraved areas (LEAs) are considered \citet{afte-article1992}.
Bullet lands are depressed areas between the grooves made by the rifling
action of the barrel. Compared to toolmarks made by screwdrivers
striation marks on bullets are typically much smaller, both in length
and in width. Bullets also have a curved cross-sectional topography.
Figure \ref{fig:rgl} shows us how the signature from a bullet land
(bottom) lines up with the image of the land (top) from which it was
extracted. We can also see in the figure how the depth and relative
position of the striation markings seen in the image are interpreted as
the signature.

Bullet matching methods are usually based on these associated
signatures. \citet{chu2013} use an automatic method for counting
consecutive matching striae (CMS). The authors report an error rate of
52\% of the known same source lands comparisons as misidentified (false
negative) and zero false positives for known different source lands.
\citet{ma2004} and \citet{vorburger2011} discuss CCF (cross-correlation
function) and its discriminating power and applicability for same-source
analyses of bullets, but do not provide any error rates in their
discussion. \citet{aoas} use multiple features like CCF, CMS, D
(distance measure) etc in a random forest based method and compare every
land against every other land of digitised versions of Hamby 252 and
Hamby 44 \citep{hamby} published on the NIST Ballistics Database
\citep{nist}. The authors report an out-of-bag overall error rate of
0.46\%, comprised of an error rate of 30.05\% of same-source pairs that
were not identified and an error rate of 0.026\% of different-source
pairs that were incorrectly identified as same-source.

The Chumbley score provides us with another approach in the same-source
assessment of bullet striation marks. \citet{chumbley} compare two
toolmarks for same-source. The data for this study was obtained from 50
sequentially manufactured screwdriver tips. \citet{chumbley} report
error rates for markings made by the tips at different angles. For
markings made at 30 degree the authors report an average false negative
error rate of 0.089 and an average false positive error rate of 0.023.
For other angles of 60 and 85 degrees the false negatives error rate is
0.09 while the rate of false positives decreases to 0.01. The paper by
\citet{hadler} is based on the same data but the authors focus on
markings made under the same angle. The error rates associated with the
deterministic version of the score are 0.06 for false negatives and a
false positive error rate of 0.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{images/B6-B2-L6-rescaled.png}

\caption{\label{fig:rgl} Image of a bullet land from a confocal light microscope at 20 fold magnification (top) and a chart of the corresponding signature of the same land (bottom). The dotted lines connect some peaks visible in both visualizations.}

\end{figure}

\hypertarget{scans-for-land-engraved-areas}{%
\subsection{Scans for land engraved
areas}\label{scans-for-land-engraved-areas}}

Comparisons of striae from bullets are usually based on comparisons of
striae in land engraved areas, which are extracted in form of cross
sections, called \emph{profiles} \citep{aoas,ma2004}. From profiles
bullet \emph{signatures} \citep{chu2013,aoas} are extracted as residuals
of a loess fit or Gaussian filter. This effectively removes topographic
structure from the data in the attempt to increase the signal to noise
ratio. The span of the loess fit was found using cross-validation, as
described by \citet{aoas}.

There are two sources of scans for sets from the Hamby study available
to us: scans of Hamby 44 and Hamby 252 are available from the NIST
database \citep{nist}. Hamby 44 has also been made available to us and
has been scanned locally for CSAFE at the Roy J.~Carver High Resolution
Microscopy Facility using a Sensofar confocal light microscope. Scans in
the NIST database are made with a NanoFocus at 20x magnification. The
resolutions of the two instruments are different: the NIST scans are
taken at a resolution of 1.5625 \(\mu m\) per pixel, while the CSAFE
scans are available at a resolution of 0.645 \(\mu m\) per pixel. The
length of an average bullet land from Hamby (9 mm Ruger P85) is about 2
millimeter, resulting in signatures of about 1200 pixels for NIST scans,
and about 3000 pixels for CSAFE scans.

In comparison, scans from the profilometer used by
\citet{chumbley, hadler} were taken at a resolution of about 0.73
\(\mu m\) per pixel. The screw driver toolmarks are about 7 mm in length
\citep{manytoolmarks1}, for a total of over 9000 pixels for the width of
these scans.

This severe limitation in the amount of available data poses the main
challenge in adapting the Chumbley score to matching bullet lands,
because of the resulting loss in power.

\hypertarget{the-chumbley-score-test}{%
\subsection{The Chumbley Score Test}\label{the-chumbley-score-test}}

The Chumbley score algorithm takes input in form of two digitized
toolmarks. The toolmark is in form of \(z(t)\) which is a spatial
process for location indexed by \(t\). \(t\) here denotes equally spaced
pixel locations for the striation marks under consideration,
\(t = 1, ..., T\). Let further \(z^s(t)\) denote a vector of markings of
length \(s\) starting in location \(t\).

Let \(x(t_1)\), \(t_1 = 1,2,...T_1\) and \(y(t_2)\), \(t_2 = 1,2...T_2\)
be two digitized toolmarks (where \(T_1\) and \(T_2\) are not
necessarily equal). The toolmarks under consideration are potentially
from two different sources or the same source. \(T_1\) and \(T_2\), as
represented above, are the final pixel indexes of each marking and
therefore give the respective lengths of the markings.

In a pre-processing step the two markings are smoothed using a lowess
\citep{lowess} with coarseness parameter \(c\). Originally, this
smoothing is intended to remove drift and (sub)class characteristics
from individual markings, however, in the setting of matching bullet
striae, we can also make use of this mechanism to separate bullet
curvature in profiles from signatures before matching signatures.

After removing sub-class structure, the Chumbley scores is calculated in
two steps: an optimization step and a validation step. In the
optimization step, the two markings are aligned horizontally such that
within a pre-defined window of length \(w_o\) the correlation between
\(x(t_1)\) and \(y(t_2)\) is maximized: \[
\left(t_1^o, t_2^o\right) = \mathop{\arg \max}\limits_{1 \le t_1 \le T_1, 1 \le t_2 \le T_2} \text{cor} \left(x^{w_o} (t_1), y^{w_o}(t_2) \right)
\] This results in an optimal vertical (in-phase) shift of
\(t_1^o - t_2^o\) for aligning the two markings.
{\textcolor{orange}{We will denote the relative optimal locations as $t_1^*$ and $t_2^*$, where $t_i^* = t_i^o/(T_i-w_o)$ for $i=1,2$, such that $t_1^*, t_2^* \in [0,1]$. Once (sub-)class characteristics are removed, the relative optimal locations should be distributed according to a uniform distribution in $[0,1]$. }}

In the validation step, two sets of windows of size \(w_v\) are chosen
from both markings (see Figure \ref{fig:win-comparison}). In the first
set, pairs of windows are extracted from the two markings using the
optimal vertical shift as determined in the first step, whereas for the
second set the windows are extracted using a different (out-of-phase)
shift.

More precisely, let us define starting points \(s_i^{(k)}\) for each
signature \(k = 1, 2\) as \begin{eqnarray}\label{eq.start}
s^{(k)}_i = 
\begin{cases}
t_k^* + i w_v & \text{ for } i < 0 \\
t_k^* + w_ o + i w_v & \text{ for } i \ge 0,
\end{cases}
\end{eqnarray} for integer values of \(i\) with
\(0 < s^{(k)}_i \le T_k - w_v\).

Same-shift pairs of length \(w_v\) are defined in \citet{hadler} as all
pairs \((s_i^{(1)}, s_i^{(2)})\) for all integer values \(i\) for which
both \(s_i^{(1)}\) and \(s_i^{(2)}\) are defined. Similarly,
different-shift pairs are defined as \((s_i^{(1)}, s_{-i-1}^{(2)})\) for
all \(i\) where both \(s_i^{(1)}\) and \(s_{-i-1}^{(2)}\) are defined
(see \autoref{sketch-same-diff}).

\begin{figure}[hbtp]
\centering
\includegraphics[width=.7\textwidth]{images/sketch-same.png}

\includegraphics[width=.7\textwidth]{images/sketch-diff.png}
\caption{\label{sketch-same-diff}Sketch of same-shift pairings  (top) and different-shift pairings (bottom). Filled in rectangles show pairings resulting in correlations, unfilled rectangles are segments without a match.}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=\textwidth]{figures/win-comparison-1}

}

\caption{Two markings made by the same source. For convenience, the markings are moved into phase on the left and out-of phase on the right. In-phase (left) and out-of-phase (right) samples are shown by the light grey background. The Chumbley-score is based on a Mann-Whitney U test of the correlations derived from these two sets of samples.}\label{fig:win-comparison}
\end{figure}

For both same- and different-shift pairs correlations between the
markings are calculated. The intuition here is that for two markings
from the same source the correlation for the in-phase sample should be
high, while the correlations of the out-of-phase sample provide a
measure for the base-level correlation for non-matching marks of a given
length \(w_v\). The Chumbley score is then computed as a Mann Whitney U
statistic to compare between in-phase sample and out-of-phase sample. In
the original method proposed in \citet{chumbley} both in-phase and
out-of-phase sample are extracted randomly, whereas \citet{hadler}
proposed the above specified deterministic rules for both samples to
make the resulting score deterministic while simultaneously avoiding
overlaps within selected marks to ensure independence.

\hypertarget{a-problem-with-failed-tests}{%
\subsection{A problem with failed
tests}\label{a-problem-with-failed-tests}}

Looking closer at \autoref{eq.start}, we see that by definition, some
number of tests will fail to produce a result.
{\textcolor{orange}{The problem of failed tests is first mentioned in \citet{afte-chumbley}. Unfortunately, the authors do not provide any percentage of how many tests failed for their data.}}
The algorithm fails to produce for two reasons: either the number of
eligible same-shift pairs is zero, or the number of different-shift
pairs is zero.

The number of same-shift pairs will be zero, if the optimal locations
\(t_1^{o}\) and \(t_2^{o}\) are so far apart, that no segments of size
\(w_v\) are left on the same sides of the optimal locations, i.e.
\(t_1^{o} < w_v\) and \(t_2^{o} > T_2-w_o-w_v\) or
\(t_1^{o} < T_1- w_o - w_v\) and \(t_2^{o} < w_v\), i.e.~we have a
failure rate of \[
P\left( t_1^{o} < w_v \ \cap \ t_2^{o} > T_2-w_o-w_v\right) + P\left( t_1^{o} < T_1- w_o - w_v \ \cap \ t_2^{o} < w_v\right).
\] While we can assume that once (sub-)class characteristics are
removed, optimal locations \(t_i^{o}\) are uniformly distributed across
the length of the profile, we cannot assume that \(t_1^o\) and \(t_2^o\)
are independent of each other. In particular, for same-source profiles,
we would expect a strong dependency between these locations, in which
case a large difference between locations is unlikely. However, for
different source matches, we can assume that locations are independent.
In that case, we expect a test to fail with a probability of
\(\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}\). For an average length of \(T_i\)
of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels this
probability is about 0.0015.

The number of possible different-shift pairs also depends on the
location of the optimal locations \(t_1^o\) and \(t_2^o\). Whenever the
optimal locations are close to the boundaries, the number of possible
pairings decreases and reaches zero, if \(t_i^{(o)} < w_v\) or
\(t_i^{(o)} > T_i-w_o- w_v\). Assuming a correlation between optimal
locations \(t_1^o\) and \(t_2^o\) of close to one for same-source
profiles, this results in an expected rate of failure of
\(2 w_v / (T_i-w_o)\), or about 5.6\% for an average length of \(T_i\)
of 1200 pixels, \(w_o = 120\) pixels and \(w_v = 30\) pixels. Assuming
independence in the optimal locations for different source profiles the
expected probability for a failed test is, again,
\(\frac{2 w_v^2}{(T_1-w_o)(T_2-w_o)}\).

\hypertarget{a-modified-approach}{%
\subsection{A modified approach}\label{a-modified-approach}}

While failures due to missing correlations from same-shift pairs are
unavoidable by definition of the Chumbley score, failures due to missing
correlations from different-shift pairs can be prevented by using a
different strategy in assigning pairs.

Using the same notation as in \autoref{eq.start}, we define same-shift
pairs identical to \citet{hadler} as pairs \((s_i^{(1)}, s_i^{(2)})\)
for all \(i\) where the boundary conditions of both sequences are met
simultaneously. Let us assume that this results in \(I\) pairs. Define
\(s_{(j)}^{(k)}\) to be the \(j\)th starting location in sequence
\(k = 1, 2\),
i.e.~\(s_{(1)}^{(k)} < s_{(2)}^{(k)} < ... < s_{(I)}^{(k)}\).

We then define the pairs for different-shifts as
\begin{equation}\label{eq.diff2}
\left(s_{(j)}^{(1)}, s_{(I-j+1)}^{(1)} \right) \text{ for } j = 
\begin{cases}
1, ..., I & \text{ for even } I \\
1, ..., (I-1)/2, (I-1)/2 + 2, ..., I & \text{ for odd } I
\end{cases},
\end{equation} i.e.~for an odd number of same-shift correlations, we
skip the middle pair for the different-shift correlations (see
\autoref{sketch-diff-2}). This pairing ensures that the number of
different-shift pairings is the same or at most one less than the number
of same-shift pairings in all tests.
{\textcolor{orange}{In the remainder of the paper, we will refer to the algorithm defined by \citet{hadler} as{\bf (CS1)} and the suggested modified algorithm as{\bf (CS2)} and compare their performance on the available scans of the Hamby study.}}

\begin{figure}[hbtp]
\centering
\includegraphics[width=.7\textwidth]{images/sketch-diff-2.png}
\caption{\label{sketch-diff-2}Sketch of adjusted different-shift pairings. At most one of the same-shift pairings can not be matched with a different-shift pair. }
\end{figure}

\begin{figure}[hbtp]
\centering
\includegraphics[width=0.62\textwidth]{images/same-shift-failure.png}

\caption{\label{same-shift-failure}Sketch of same-shift pairings  (top) when the lag is too large to accomodate a a vaildation window in either of the two signatures}
\end{figure}

\begin{figure}[hbtp]
\centering
\includegraphics[width=0.7\textwidth]{images/diff-shift-failure.png}

\caption{\label{diff-shift-failure}Sketch of diff-shift pairings  (top) when the number of diff-shift computations is likely to 0}
\end{figure}

\begin{figure}[hbtp]
\centering
\includegraphics[width=.7\textwidth]{images/CS1-failure-CS2-success.png}
\caption{\label{CS1-failure-CS2-success}Sketch of a case where CS1 fails but CS2 does not fail}
\end{figure}

\hypertarget{testing-setup}{%
\section{Testing setup}\label{testing-setup}}

\hypertarget{the-data}{%
\subsection{The Data}\label{the-data}}

\begin{figure}

{\centering \includegraphics[width=\textwidth]{figures/sigs-profiles-1}

}

\caption{Bullet land profile (left) and the corresponding signature (right) for one of the lands of Hamby-44.}\label{fig:sigs-profiles}
\end{figure}

Lands for all Hamby-44 and Hamby-252 scans are made available through
the NIST ballistics database \citep{nist} and are considered, here. Both
of these sets of scans are part of the larger Hamby study \citep{hamby}.
Each set consists of twenty known bullets (two each from ten
consecutively rifled Ruger P85 barrels) and fifteen questioned bullets
(each matching one of the ten barrels). Ground truth for both of these
Hamby sets is known and was used to assess correctness of the tests
results.

Profiles for each bullet land were extracted from scans close to the
heel of the bullet while avoiding break-off as described in
\citet{aoas}.

\hypertarget{setup}{%
\subsection{Setup}\label{setup}}

Both algorithms (CS1) and (CS2) are implemented in the R package
\texttt{toolmaRk} \citep{toolmark}. We applied both methods to all
pairwise land-to-land comparisons of the Hamby scans provided by NIST
for a total of 85,491 land-to-land comparisons.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{failed-tests}{%
\subsection{Failed Tests}\label{failed-tests}}

Initially, the default settings suggested in \citet{hadler} were used:
\(w_o = 120\) pixels or about 190 \(\mu m\) (ten percent of the average
length of profiles) and coarseness \(c = 0.25\). \autoref{fig:fails}
shows the percentage of failed tests among the 85,491 land-to-land
comparisons of the NIST data for different values of the validation
window size \(w_v\). For same-source lands up to 12.5 percent of the
tests fail using CS1. The highest percentage of failed tests under CS2
is 1.3\% for different source tests using a validation window size
\(w_v\) of 60 pixels.
{\textcolor{orange}{Rates of expected failures are based on simulation runs using covariances between locations of same-source profiles of 0.854, and 0.120 for locations from different-source profiles, matching observed covariances for the Hamby scans. }}{\textcolor{orange}{ Observed failure rates are higher than expected. This might be due to remaining sub-class structure at a coarseness of 0.3 resulting in a distribution of optimal locations different from the assumed uniform. }}

\begin{figure}

{\centering \includegraphics[width=0.7\textwidth]{figures/fails-1}

}

\caption{Percent of failed land-to-land comparisons using an optimization window $w_o = 120$ and a coarseness of $c = 0.25$. With an increase in the size of the validation window  a higher percentage of tests fails under both methods (CS1) and (CS2), but the percentage of failed tests is much smaller under (CS2). Observed failure rates are higher than expected rates.}\label{fig:fails}
\end{figure}

\hypertarget{coarseness}{%
\subsection{Coarseness}\label{coarseness}}

The purpose of the coarseness parameter is to remove (sub-)class
characteristics from profiles before comparisons for matching.
\citet{hadler} suggest a coarseness parameter of 0.25 in the setting of
toolmark comparisons. For bullet lands, coarseness might need to be
adjusted because of the strong effect bullet curvature has on profiles.

{\textcolor{orange}{\autoref{fig:profile-sketch} gives an overview of the effect of different coarseness parameters: from left to right, coarseness levels $c$ are varied in steps of 0.05 from 0.1 to 0.3. The top row shows  resulting signature after smoothing the profile shown in Figure~\ref{fig:sigs-profiles} with the coarseness specified. 
The histograms in the bottom row show the relative optimal location $t^*$. Optimal locations are distributed uniformly once (sub-)class characteristics are removed. However, for coarseness values of $c > 0.20$ we see quite distinct boundary effects: optimal locations $t^*$ are found at the very extreme ends of a profile more often than one would expect based on a uniform distribution.}}

{\textcolor{blue}{The key effect of the optimal locations and thereby the coarseness is seen in the number of failed tests. Irrespective of whether CS1 or CS2 is being used, if the relative optimal locations are at the boundaries we will see an increase in the number of failed tests. A balance is therefore needed in the selection of the coarseness parameter which reduces the boundary effect but also does not remove important (sub-)class characteristics}}

\begin{figure}

{\centering \includegraphics[width=\textwidth]{figures/profile-sketch-1}

}

\caption{Overview of the effect of different coarseness parameters $c$ on the profile shown in Figure \ref{fig:sigs-profiles} (top). The bottom row shows histograms of  the (relative) optimal locations $t^o$ identified in the optimization step for different values of the coarseness parameter $c$. }\label{fig:profile-sketch}
\end{figure}

\hypertarget{type-ii-error-rates}{%
\subsection{Type II error rates}\label{type-ii-error-rates}}

\autoref{fig:type2} gives an overview of the type 2 error of methods CS1
and CS2 across a range of different optimization windows \(w_o\).

\begin{figure}

{\centering \includegraphics[width=0.9\textwidth]{figures/type2-1}

}

\caption{Type II error rates observed across a range of window sizes for optimization $w_o$. For a window size of $w_o = 130$ we see a minimum in type II error rate across all type I rates considered. Smaller validation sizes $w_v$ are typically associated with a smaller type II error.}\label{fig:type2}
\end{figure}

\autoref{fig:roc} gives an overview of ROC (Receiver operating
characteristic) curves for methods CS1 and CS2 over a range of different
optimization window sizes \(w_o\). The different color hues represent
the two methods CS1 (red) and CS2 (blue). The ROC curves show the
superior performance of CS2 over CS1. Generally, an optimization window
\(w_o\) of 150 pixels or more leads to the best performance with respect
to ROC curves.

\begin{figure}

{\centering \includegraphics[width=\textwidth]{figures/roc-1}

}

\caption{ROC curves of methods CS1 and CS2 for different sizes of optimization window $w_o$. Best performances with respect to  ROC curves are reached for optimization windows of sizes 150 and higher. Points of equal error rates (EERs) can be found at the intersection of the dotted line and the ROC curves.}\label{fig:roc}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=\textwidth]{figures/eer-1}

}

\caption{Comparison of results for CS1 and CS2 using area under the curve (AUC) and equal error rates (EER). }\label{fig:eer}
\end{figure}

\hypertarget{observed-versus-nominal-type-i-error-rates}{%
\subsection{Observed versus Nominal Type I error
rates}\label{observed-versus-nominal-type-i-error-rates}}

\autoref{fig:type1} shows observed type I error rates across a range of
optimization windows \(w_o\). Generally, observed type I errors are
higher than expected. Method CS1 shows in this instance slightly better
performance than method CS2, but for both an increase in the size of the
optimization window leads to a decrease in the observed type I error
rates.

\begin{figure}

{\centering \includegraphics[width=\textwidth]{figures/type1-1}

}

\caption{Comparison of observed and nominal type I error rates  across a range of window sizes for optimization $wo$. The horizontal line in each facet indicates the nominal type I error rate.}\label{fig:type1}
\end{figure}

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

{\textcolor{orange}{
We started out trying to assess the suitability of the Chumbley Score's suitability for matching striae on bullet lands. Besides using recommended defaults, we also proposed ways to optimize parameters: based on the assumption that once sub-class characteristics are removed, optimal locations are distributed uniformly across the profile, we found  for bullets the smaller coarseness value of $c = 0.15$ to be suitable. 
Optimized window sizes were found based on cross-validation to minimize type 2 error rates. 
Method CS1 proposed by \citet{hadler} has a minimal type 2 error rate of XXX for an optimized window size of YYY - which is considerably higher than the error rates achieved on matching toolmarks. 
Unfortunately, method CS1 also has a high rate of failed tests -- situations, in which the algorithm does not provide a result, due to the way different-shift pairs are constructed. 
XXX
The algorithm CS2 introduced here is constructed in a way that reduces the number of failures by a magnitude of about XXX. While reducing the failure rate, the algorithm also shows an increase in the power of the test. Type II error rates reach a minimum of XXX for an optimized window size of YYY. While significantly reduced over CS1, CS2 still has type 2 error rates on bullet lands that are  higher than the error rates achieved on the -much larger- toolmarks.
However, bullets usually have multiple lands - in the case of Ruger P85s as used in the Hamby study, there are six lands for each bullet. We might be able to get more power out of the test by adapting CS2 to deal work in a bullet-to-bullet comparison.
}}

{\textcolor{orange}{
Fix in CS2 should also improve power for matching toolmarks. 
}}

\bibliographystyle{agsm}
\bibliography{bibliography}

\end{document}